██╗   ██╗██████╗ ███████╗███████╗████████╗    ██████╗ ██╗      ██████╗ ████████╗
██║   ██║██╔══██╗██╔════╝██╔════╝╚══██╔══╝    ██╔══██╗██║     ██╔═══██╗╚══██╔══╝
██║   ██║██████╔╝███████╗█████╗     ██║       ██████╔╝██║     ██║   ██║   ██║   
██║   ██║██╔═══╝ ╚════██║██╔══╝     ██║       ██╔═══╝ ██║     ██║   ██║   ██║   
╚██████╔╝██║     ███████║███████╗   ██║       ██║     ███████╗╚██████╔╝   ██║   
 ╚═════╝ ╚═╝     ╚══════╝╚══════╝   ╚═╝       ╚═╝     ╚══════╝ ╚═════╝    ╚═╝   
                                                                                
# Prepare environment
```{bash echo=FALSE, engine.opts='-l'}
# Remove old results
rm -r .temp/

# Create directory tree
mkdir .temp/

# Move to working dir
cd .temp/

# Set paths
utils=../../Utils
H_catalogs=../../Data/H_catalogs/

# Process lncRNA catalogues <- exclude chrM, exclude unspliced
zcat $H_catalogs/NONCODE.gtf.gz | awk 'BEGIN{FS=OFS="\t"}$7!="."' | awk '$1 ~ /^chr[0-9XY]{1,2}$/ {print $0}' | $utils/sortgff - | $utils/gff2bed_full.pl - | awk '$10 > 1 {print $0}' | $utils/bed12togff - | grep "\S" > NONCODE.hg38.gtf
zcat $H_catalogs/refSeq.gtf.gz | awk 'BEGIN{FS=OFS="\t"}$7!="."' | awk '$1 ~ /^chr[0-9XY]{1,2}$/ {print $0}' | $utils/sortgff - | $utils/gff2bed_full.pl - | awk '$10 > 1 {print $0}' | $utils/bed12togff - | grep "\S" > refSeq.hg38.gtf
zcat $H_catalogs/miTranscriptome.gtf.gz | awk 'BEGIN{FS=OFS="\t"}$7!="."' | awk '$1 ~ /^chr[0-9XY]{1,2}$/ {print $0}' | $utils/sortgff - | $utils/gff2bed_full.pl - | awk '$10 > 1 {print $0}' | $utils/bed12togff - | grep "\S" > miTranscriptome.hg38.gtf
zcat $H_catalogs/gencode20+.gtf.gz | awk 'BEGIN{FS=OFS="\t"}$7!="."' | awk '$1 ~ /^chr[0-9XY]{1,2}$/ {print $0}' | $utils/sortgff - | $utils/gff2bed_full.pl - | awk '$10 > 1 {print $0}' | $utils/bed12togff - | grep "\S" > gencode20+.hg38.gtf
zcat $H_catalogs/bigTranscriptome.gtf.gz | awk 'BEGIN{FS=OFS="\t"}$7!="."' | awk '$1 ~ /^chr[0-9XY]{1,2}$/ {print $0}' | $utils/sortgff - | $utils/gff2bed_full.pl - | awk '$10 > 1 {print $0}' | $utils/bed12togff - | grep "\S" > bigTranscriptome.hg38.gtf
zcat $H_catalogs/fantomCat.gtf.gz | awk 'BEGIN{FS=OFS="\t"}$7!="."' | awk '$1 ~ /^chr[0-9XY]{1,2}$/ {print $0}' | $utils/sortgff - | $utils/gff2bed_full.pl - | awk '$10 > 1 {print $0}' | $utils/bed12togff - | grep "\S" > fantomCat.hg38.gtf
# Add chess and lncBook (convert to bed12 and then gtf to remove unnecessary data)
zcat $H_catalogs/chess3_1_2_GRCh38_lncRNA.gtf.gz | $utils/gff2bed_full.pl - | awk 'BEGIN{FS=OFS="\t"}$6!="."' | awk '$1 ~ /^chr[0-9XY]{1,2}$/ {print $0}' | awk '$10 > 1 {print $0}' | grep "\S" | $utils/bed12togff - > chess3_1_2.hg38.gtf
zcat $H_catalogs/lncRNA_LncBookv2_0_GRCh38.gtf.gz | $utils/gff2bed_full.pl - | awk 'BEGIN{FS=OFS="\t"}$6!="."' | awk '$1 ~ /^chr[0-9XY]{1,2}$/ {print $0}' | awk '$10 > 1 {print $0}' | grep "\S" | $utils/bed12togff - > lncBookv2_0.hg38.gtf
# Add gencode 27 and 47
## Extract spliced transcriptIDs - v47
zcat $H_catalogs/gen47.gtf.gz | awk 'BEGIN{FS=OFS="\t"}$7!="."' | awk '$1 ~ /^chr[0-9XY]{1,2}$/ {print $0}' | awk '$3=="exon"' | grep "\S" | $utils/sortgff - | $utils/gff2bed_full.pl - | awk '$10 > 1 {print $4}' | sort | uniq | grep "\S" > gen47_spliced_tx.ids
## Extract clean v47 slice
zcat $H_catalogs/gen47.gtf.gz | awk 'BEGIN{FS=OFS="\t"}$7!="."' | awk '$1 ~ /^chr[0-9XY]{1,2}$/ {print $0}' | awk '$3=="exon"' | grep -Fwf gen47_spliced_tx.ids | grep "\S" > gen47.hg38.gtf
## Extract spliced transcriptIDs - v47
zcat $H_catalogs/gen27.gtf.gz | awk 'BEGIN{FS=OFS="\t"}$7!="."' | awk '$1 ~ /^chr[0-9XY]{1,2}$/ {print $0}' | awk '$3=="exon"' | grep "\S" | $utils/sortgff - | $utils/gff2bed_full.pl - | awk '$10 > 1 {print $4}' | sort | uniq | grep "\S" > gen27_spliced_tx.ids
## Extract clean v27 slice
zcat $H_catalogs/gen27.gtf.gz | awk 'BEGIN{FS=OFS="\t"}$7!="."' | awk '$1 ~ /^chr[0-9XY]{1,2}$/ {print $0}' | awk '$3=="exon"' | grep -Fwf gen27_spliced_tx.ids | grep "\S" > gen27.hg38.gtf

# Include concatenated catalogs
## Concatenate human catalogs into 1 file (ENST00000610119.1 has inconsistent strand); remove chrM
cat NONCODE.hg38.gtf refSeq.hg38.gtf miTranscriptome.hg38.gtf gencode20+.hg38.gtf bigTranscriptome.hg38.gtf fantomCat.hg38.gtf | $utils/gff2bed_full.pl - | sort | uniq | grep -vFw "ENST00000610119.1" | awk 'BEGIN{FS=OFS="\t"}$6!="."' | awk '$1 ~ /^chr[0-9XY]{1,2}$/ {print $0}' | grep "\S" | awk '$10 > 1 {print $0}' | $utils/bed12togff - | $utils/sortgff - | $utils/tmerge --exonOverhangTolerance 8 --tmPrefix IC - | grep '\S' > H_concat_catalogs.hg38.gtf

# Prepare config file
## Ends
echo "5" >> ends.config
echo "3" >> ends.config

# Prepare Config files
for catalogue in `ls *.hg38.gtf`; do echo $catalogue >> upSet.config; done

# Run buildLoci for all except gencode 27 and gencode 47
echo "Running buildLoci"
while read file || [ -n "$file" ]
  do
      name=`echo $file | awk -F'.' '{print $1}'`
      echo "Name: $name"
      if [[ "$name" == "gen27" ]] || [[ "$name" == "gen47" ]]; then
         $utils/gff2gff.pl $name.hg38.gtf > $name.hg38.gene.gtf
         $utils/extract_locus_coords.pl $name.hg38.gene.gtf > $name.hg38.gene.bed
      
      else
         bedtools intersect -s -wao -a $file -b $file | $utils/buildLoci.pl - > $name.gene.tmp.gtf
         $utils/gff2gff.pl $name.gene.tmp.gtf > $name.hg38.gene.gtf
         $utils/extract_locus_coords.pl $name.hg38.gene.gtf > $name.hg38.gene.bed
      fi
  done < ./upSet.config

# Report completeness
echo "######### COMPLETED"
```

# Generate data
```{bash echo=FALSE, engine.opts='-l'}
# Move to working dir
cd .temp/

# Remove files from previous run
rm -r upSet/

# Create sub-dir to separate files and go there
mkdir upSet/ && cd upSet/

# Set paths
utils=../../../Utils

# Handle errors
set -e          # exit on any non-0 exit status
set -o pipefail # exit on any non-0 exit status in pipe

# Create master bed by concatenating catalogs
echo "++ Concatenating gene files..."
while read file || [ -n "$file" ]
  do
    name=`echo $file | awk -F'.' '{print $1}'`
    # Include catalog name in geneID (otherwise we will end up with duplicated buildLoci ids)
    cat ../$name.hg38.gene.bed | awk -F'\t' -v name="$name" 'BEGIN {OFS=FS} {$4=name "|" $4; print $0}' >> master.gene.temp.bed
  done < ../upSet.config
echo ">> Done"

## Run bedtools intersect to find genes that are intersecting each other
echo "++ Running bedtools intersect to find mutually intersectin genes..."
bedtools intersect -s -wa -wb -f 1.00 -F 1.00 -e -a master.gene.temp.bed -b master.gene.temp.bed > master.intersecting.master.bed
echo ">> Done"
## Extract geneIDs that are not intersecting with each other
echo "++ Extracting IDs of genes, that are not intersecting each other..."
$utils/removeDuplicatedGenes.py master.intersecting.master.bed > master.gene.ids
echo ">> Done"
## Remove intersecting genes
echo "++ Removing intersecting genes from master.gene.temp.bed file..."
cat master.gene.temp.bed | grep -Fwf master.gene.ids > master.gene.bed
echo ">> Done"

# Remove duplicates from concatenated master.temp.bed
#echo "++ Removing duplicated genes..."
#cat master.gene.temp.bed | awk '!seen[$1,$2,$3,$6]++{print $0}' > master.gene.bed
#echo ">> Done"

# Extract only geneIDs
#echo "++ Extracting geneIDs from master.bed ..."
#cat master.gene.bed | awk '{print $4}' > master.gene.ids
#echo ">> Done"

# Run intersecting
echo "++ Running intersecting... "

while read file || [ -n "$file" ]
  do
    name=`echo $file | awk -F'.' '{print $1}'`
    echo "~~ Intersecting: $name"
    bedtools intersect -s -wa -wb -f 1.00 -F 1.00 -e -a master.gene.bed -b ../$name.hg38.gene.bed > master.intersecting.$name.bed
    echo ">> Done"
    
    echo "~~ Extracting genes..."
    cat master.intersecting.$name.bed | awk '!seen[$4]++{print $4}' > master.intersecting.$name.gene.ids
    echo ">> Done"
    
  done < ../upSet.config

# Report completeness
echo "######### COMPLETED"
```

# Load libraries
```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,
               data.table,
               UpSetR)
```

# Process data and create UpSet plot
```{r}
# Clean the env
rm(list = ls())

# Load data
df_masterIDs <- read.table(".temp/upSet/master.gene.ids", header=F, sep="\t") %>%
  rename(geneID = "V1") %>% 
  mutate(master = 1)

df_inter_bigtrans <- read.table(".temp/upSet/master.intersecting.bigTranscriptome.gene.ids", header=F, sep="\t") %>%
  rename(geneID = "V1") %>% 
  mutate(bigtrans = 1)
df_inter_chess3_1_2 <- read.table(".temp/upSet/master.intersecting.chess3_1_2.gene.ids", header=F, sep="\t") %>%
  rename(geneID = "V1") %>% 
  mutate(chess3_1_2 = 1)
df_inter_fantomCat <- read.table(".temp/upSet/master.intersecting.fantomCat.gene.ids", header=F, sep="\t") %>%
  rename(geneID = "V1") %>% 
  mutate(fantomCat = 1)
df_inter_gen27 <- read.table(".temp/upSet/master.intersecting.gen27.gene.ids", header=F, sep="\t") %>%
  rename(geneID = "V1") %>% 
  mutate(gen27 = 1)
df_inter_gen47 <- read.table(".temp/upSet/master.intersecting.gen47.gene.ids", header=F, sep="\t") %>%
  rename(geneID = "V1") %>% 
  mutate(gen47 = 1)
df_inter_gencodeLncRna <- read.table(".temp/upSet/master.intersecting.gencode20+.gene.ids", header=F, sep="\t") %>%
  rename(geneID = "V1") %>% 
  mutate(gencodeLncRna = 1)
df_inter_H_concat_catalogs <- read.table(".temp/upSet/master.intersecting.H_concat_catalogs.gene.ids", header=F, sep="\t") %>%
  rename(geneID = "V1") %>% 
  mutate(H_concat_catalogs = 1)
df_inter_lncBookv2_0 <- read.table(".temp/upSet/master.intersecting.lncBookv2_0.gene.ids", header=F, sep="\t") %>%
  rename(geneID = "V1") %>% 
  mutate(lncBookv2_0 = 1)
df_inter_mitrans <- read.table(".temp/upSet/master.intersecting.miTranscriptome.gene.ids", header=F, sep="\t") %>%
  rename(geneID = "V1") %>% 
  mutate(mitrans = 1)
df_inter_noncode <- read.table(".temp/upSet/master.intersecting.NONCODE.gene.ids", header=F, sep="\t") %>%
  rename(geneID = "V1") %>% 
  mutate(noncode = 1)
df_inter_refseq <- read.table(".temp/upSet/master.intersecting.refSeq.gene.ids", header=F, sep="\t") %>%
  rename(geneID = "V1") %>% 
  mutate(refseq = 1)

# Join tables
df_masterIDs %>% 
  left_join(x = .,
            y = df_inter_bigtrans,
            by = "geneID") %>% 
  left_join(x = .,
            y = df_inter_chess3_1_2,
            by = "geneID") %>% 
  left_join(x = .,
            y = df_inter_fantomCat,
            by = "geneID") %>% 
  left_join(x = .,
            y = df_inter_gen27,
            by = "geneID") %>% 
  left_join(x = .,
            y = df_inter_gen47,
            by = "geneID") %>% 
  left_join(x = .,
            y = df_inter_gencodeLncRna,
            by = "geneID") %>% 
  left_join(x = .,
            y = df_inter_H_concat_catalogs,
            by = "geneID") %>% 
  left_join(x = .,
            y = df_inter_lncBookv2_0,
            by = "geneID") %>% 
  left_join(x = .,
            y = df_inter_mitrans,
            by = "geneID") %>% 
  left_join(x = .,
            y = df_inter_noncode,
            by = "geneID") %>% 
  left_join(x = .,
            y = df_inter_refseq,
            by = "geneID") %>% 
  # Exclude master column
  select(!master) %>% 
  # Convert NA to 0
  mutate(across(where(is.numeric), ~ifelse(is.na(.x), 0, .x))) -> df_one_hot_encoded

# Prepare labels  
pretty_labels <- c(
  chess3_1_2 = "CHESS 3.1.2",
  lncBookv2_0 = "lncBook v2.0",
  bigtrans = "bigTranscriptome",
  gencodeLncRna = "GENCODEv20+",
  refseq = "refSeq",
  fantomCat = "fantomCat",
  mitrans = "miTranscriptome",
  noncode = "NONCODE",
  H_concat_catalogs = "lncRNA-merge",
  gen27 = "GENCODE v27",
  gen47 = "GENCODE v47"
)

# Create Upset plot - with lncRNA-merge
df_one_hot_encoded %>% 
  as.data.table() %>%
  setnames(old = names(pretty_labels), new = pretty_labels) %>%
  upset(., 
        sets = pretty_labels,
        mb.ratio = c(0.55, 0.45),
        keep.order = TRUE,
        order.by = "freq",
        set_size.show = TRUE,
        set_size.numbers_size = 4) -> upSet_plot

# Preview
upSet_plot
```

# Save plot to pdf - with lncRNA-merge
```{r}
pdf("upSet_plot.pdf", bg = "white", width=14, height=9)
upSet_plot
```

######################################################################################
######################################################################################
######################################################################################
# Generate data for alternative version of upSet plot using intron-chains
# Plan:
#      - All catalogs including GENCODE will be filtered (main chr, nothing else) 
#      - All catalogs including GENCODE will be tmerged to reduce redundancy
#      - All catalogs including GENCODE will be concatenated and duplicated bed12 entires will be filtered out
#      - Using grep I will find overlapping intron-chains
```{bash echo=FALSE, engine.opts='-l'}
# Remove old results
rm -r .temp2/

# Create directory tree
mkdir .temp2/

# Move to working dir
cd .temp2/

# Set paths
utils=../../Utils
H_catalogs=../../Data/H_catalogs/

# Handle errors
set -e          # exit on any non-0 exit status
set -o pipefail # exit on any non-0 exit status in pipe

# Extract non-redundant intron-chains from each lncRNA catalog
echo "++ Transforming catalogs... "
for catalog in bigTranscriptome fantomCat gencode20+ miTranscriptome NONCODE refSeq chess3_1_2_GRCh38_lncRNA lncRNA_LncBookv2_0_GRCh38 gen27 gen47
  do
    echo -e "++ Transforming ${catalog}... "
    zcat $H_catalogs/$catalog.gtf.gz | $utils/gff2bed_full.pl - | sort | uniq | awk '$1 ~ /^chr[0-9XY]{1,2}$/ {print $0}' > $catalog.bed12
    cat $catalog.bed12 | $utils/bed12togff - | $utils/sortgff - | $utils/tmerge --exonOverhangTolerance 8 --tmPrefix IC - | $utils/gff2bed_full.pl - | grep '\S' > $catalog.merged.bed12
   echo ">> Done"
  done

# Create merged catalog 
echo "++ Creating marged catalog... "
cat bigTranscriptome.bed12 fantomCat.bed12 gencode20+.bed12 miTranscriptome.bed12 NONCODE.bed12 refSeq.bed12 | sort | uniq | grep -vFw "ENST00000610119.1" | grep "\S" | grep -vFw "ENST00000610119.1" | $utils/bed12togff - | $utils/sortgff - | $utils/tmerge --exonOverhangTolerance 8 --tmPrefix IC - | grep '\S' | $utils/gff2bed_full.pl - | awk -F'\t' '!seen[$1,$2,$3,$6,$10,$11,$12]++ {print $0}' | grep '\S' > H_concat_catalogs.merged.bed12
echo ">> Done"

# Concatenate all catalogs, remove duplicated intron-chains and switch txID with a unique signature | extract only signature
echo "++ Creating master catalog... "
cat bigTranscriptome.merged.bed12 fantomCat.merged.bed12 gencode20+.merged.bed12 miTranscriptome.merged.bed12 NONCODE.merged.bed12 refSeq.merged.bed12 H_concat_catalogs.merged.bed12 chess3_1_2_GRCh38_lncRNA.merged.bed12 lncRNA_LncBookv2_0_GRCh38.merged.bed12 gen27.merged.bed12 gen47.merged.bed12 | awk -F'\t' '!seen[$1,$2,$3,$6,$10,$11,$12]++ {print $0}' | awk -F'\t' '{OFS="\t"} $4=$1"|"$2"|"$3"|"$6"|"$10"|"$11"|"$12 {print $4}' > master.IC.sig
echo ">> Done"

# Check intersection of each catalog with master
echo "++ Intersecting catalogs with master... "
for catalog in bigTranscriptome fantomCat gencode20+ miTranscriptome NONCODE refSeq H_concat_catalogs chess3_1_2_GRCh38_lncRNA lncRNA_LncBookv2_0_GRCh38 gen27 gen47
  do
    echo -e "++ Intersecting ${catalog}... "
    # Extract signature
    cat $catalog.merged.bed12 | awk -F'\t' '$4=$1"|"$2"|"$3"|"$6"|"$10"|"$11"|"$12 {print $4}' | sort | uniq | grep "\S" > $catalog.merged.sig
    # Search signature in master
    cat master.IC.sig | grep -Fwf $catalog.merged.sig > $catalog.intersecting.master.sig
    echo ">> Done"
  done

# Report completness
echo "######## COMPLETED"
```

# Load libraries
```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,
               data.table,
               UpSetR)
```

# Process data and create UpSet plot
```{r}
# Clean the env
rm(list = ls())

# Load data
df_masterIDs <- read.table(".temp2/master.IC.sig", header=F, sep="\t") %>%
  rename(icID = "V1") %>% 
  mutate(master = 1)

df_inter_bigtrans <- read.table(".temp2/bigTranscriptome.intersecting.master.sig", header=F, sep="\t") %>%
  rename(icID = "V1") %>% 
  mutate(bigtrans = 1)
df_inter_chess3_1_2 <- read.table(".temp2/chess3_1_2_GRCh38_lncRNA.intersecting.master.sig", header=F, sep="\t") %>%
 rename(icID = "V1") %>%
 mutate(chess3_1_2 = 1)
df_inter_fantomCat <- read.table(".temp2/fantomCat.intersecting.master.sig", header=F, sep="\t") %>%
  rename(icID = "V1") %>% 
  mutate(fantomCat = 1)
df_inter_gen27 <- read.table(".temp2/gen27.intersecting.master.sig", header=F, sep="\t") %>%
  rename(icID = "V1") %>% 
  mutate(gen27 = 1)
df_inter_gen47 <- read.table(".temp2/gen47.intersecting.master.sig", header=F, sep="\t") %>%
  rename(icID = "V1") %>% 
  mutate(gen47 = 1)
df_inter_gencodeLncRna <- read.table(".temp2/gencode20+.intersecting.master.sig", header=F, sep="\t") %>%
  rename(icID = "V1") %>% 
  mutate(gencodeLncRna = 1)
df_inter_H_concat_catalogs <- read.table(".temp2/H_concat_catalogs.intersecting.master.sig", header=F, sep="\t") %>%
  rename(icID = "V1") %>%
  mutate(H_concat_catalogs = 1)
df_inter_lncBookv2_0 <- read.table(".temp2/lncRNA_LncBookv2_0_GRCh38.intersecting.master.sig", header=F, sep="\t") %>%
  rename(icID = "V1") %>%
  mutate(lncBookv2_0 = 1)
df_inter_mitrans <- read.table(".temp2/miTranscriptome.intersecting.master.sig", header=F, sep="\t") %>%
  rename(icID = "V1") %>% 
  mutate(mitrans = 1)
df_inter_noncode <- read.table(".temp2/NONCODE.intersecting.master.sig", header=F, sep="\t") %>%
  rename(icID = "V1") %>% 
  mutate(noncode = 1)
df_inter_refseq <- read.table(".temp2/refSeq.intersecting.master.sig", header=F, sep="\t") %>%
  rename(icID = "V1") %>% 
  mutate(refseq = 1)

# Join tables
df_masterIDs %>% 
  left_join(x = .,
            y = df_inter_bigtrans,
            by = "icID") %>% 
  left_join(x = .,
            y = df_inter_chess3_1_2,
            by = "icID") %>%
  left_join(x = .,
            y = df_inter_fantomCat,
            by = "icID") %>% 
  left_join(x = .,
            y = df_inter_gen27,
            by = "icID") %>% 
  left_join(x = .,
            y = df_inter_gen47,
            by = "icID") %>% 
  left_join(x = .,
            y = df_inter_gencodeLncRna,
            by = "icID") %>% 
  left_join(x = .,
            y = df_inter_H_concat_catalogs,
            by = "icID") %>%
  left_join(x = .,
            y = df_inter_lncBookv2_0,
            by = "icID") %>%
  left_join(x = .,
            y = df_inter_mitrans,
            by = "icID") %>% 
  left_join(x = .,
            y = df_inter_noncode,
            by = "icID") %>% 
  left_join(x = .,
            y = df_inter_refseq,
            by = "icID") %>% 
  # Exclude master column
  select(!master) %>% 
  # Convert NA to 0
  mutate(across(where(is.numeric), ~ifelse(is.na(.x), 0, .x))) -> df_one_hot_encoded

# Prepare labels  
pretty_labels <- c(
  chess3_1_2 = "CHESS 3.1.2",
  lncBookv2_0 = "lncBook v2.0",
  bigtrans = "bigTranscriptome",
  gencodeLncRna = "GENCODEv20+",
  refseq = "refSeq",
  fantomCat = "fantomCat",
  mitrans = "miTranscriptome",
  noncode = "NONCODE",
  H_concat_catalogs = "lncRNA-merge",
  gen27 = "GENCODE v27",
  gen47 = "GENCODE v47"
)

# Create Upset plot - with lncRNA-merge
df_one_hot_encoded %>% 
  as.data.table() %>%
  setnames(old = names(pretty_labels), new = pretty_labels) %>%
  upset(., 
        sets = pretty_labels,
        mb.ratio = c(0.55, 0.45),
        keep.order = TRUE,
        order.by = "freq",
        set_size.show = TRUE,
        set_size.numbers_size = 4) -> upSet_plot

# Preview
upSet_plot
```

# Save plot to pdf - with lncRNA-merge
```{r}
pdf("IC_upSet_plot.pdf", bg = "white", width=14, height=9)
upSet_plot
```




######################################################################################
######################################################################################
######################################################################################
# Create python venv - run it once
```{bash echo=FALSE, engine.opts='-l'}
python3.11 -m venv .venv
source .venv/bin/activate
python -m pip install --upgrade pip
python -m pip install pandas networkx
```

# Generate data for alternative version of upSet plot using exonic gene structure
# Plan:
#      - Running on a prayer
```{bash echo=FALSE, engine.opts='-l'}
# Remove old results
rm -r .temp3/

# Create directory tree
mkdir .temp3/

# Move to working dir
cd .temp3/

# Set paths
utils=../../Utils
H_catalogs=../../Data/H_catalogs/

# Activate python venv
source ../.venv/bin/activate

# Handle errors
set -e          # exit on any non-0 exit status
set -o pipefail # exit on any non-0 exit status in pipe

# Transform data to bed6
echo "++ Transforming catalogs... "
for catalog in miTranscriptome NONCODE gen47 gen27 bigTranscriptome fantomCat gencode20+ refSeq chess3_1_2_GRCh38_lncRNA lncRNA_LncBookv2_0_GRCh38
  do
    echo -e "++ Transforming ${catalog}... "
    # Keep only main chromosomes | transforming from gff to bed and then to gff is to fix lncBook 9th field!
    zcat $H_catalogs/$catalog.gtf.gz | $utils/gff2bed_full.pl - | $utils/bed12togff - | awk -F'\t' '$3=="exon"' | awk '$1 ~ /^chr[0-9XY]{1,2}$/ {print $0}' | awk -F'\t' '$7!="."' > $catalog.gtf
    # Run buildLoci
    bedtools intersect -s -wao -a $catalog.gtf -b $catalog.gtf | $utils/buildLoci.pl - > $catalog.tmp.gtf
    # Fix gtf
    $utils/gff2gff.pl $catalog.tmp.gtf > $catalog.loci.gtf
    # Convert to bed6 with geneID in $4
    cat $catalog.loci.gtf | $utils/gff2bed_simplistic_geneID - | awk -F'\t' -v catalog=$catalog '{OFS=FS} $4=catalog"|"$4'  | sort | uniq  > $catalog.exons.bed6
   echo ">> Done"
  done

# Run custom script to get upSet matrix
echo "++ Running custom python script... "
python3 ../build_consensus_upset_matrix.py \
  --catalogs GENCODE47=gen47.exons.bed6 GENCODE27=gen27.exons.bed6 \
             miTranscriptome=miTranscriptome.exons.bed6 NONCODE=NONCODE.exons.bed6 \
  --min-frac 0.3 \
  --outdir results/
echo ">> Done"

# Report completness
echo "######## COMPLETED"
```

# Load libraries
```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,
               data.table,
               UpSetR)
```

# Process data and create UpSet plot
```{r}
# Clean the env
rm(list = ls())

# Load data
df_masterIDs <- read.table(".temp2/master.IC.sig", header=F, sep="\t") %>%
  rename(icID = "V1") %>% 
  mutate(master = 1)

df_inter_bigtrans <- read.table(".temp2/bigTranscriptome.intersecting.master.sig", header=F, sep="\t") %>%
  rename(icID = "V1") %>% 
  mutate(bigtrans = 1)
df_inter_chess3_1_2 <- read.table(".temp2/chess3_1_2_GRCh38_lncRNA.intersecting.master.sig", header=F, sep="\t") %>%
 rename(icID = "V1") %>%
 mutate(chess3_1_2 = 1)
df_inter_fantomCat <- read.table(".temp2/fantomCat.intersecting.master.sig", header=F, sep="\t") %>%
  rename(icID = "V1") %>% 
  mutate(fantomCat = 1)
df_inter_gen27 <- read.table(".temp2/gen27.intersecting.master.sig", header=F, sep="\t") %>%
  rename(icID = "V1") %>% 
  mutate(gen27 = 1)
df_inter_gen47 <- read.table(".temp2/gen47.intersecting.master.sig", header=F, sep="\t") %>%
  rename(icID = "V1") %>% 
  mutate(gen47 = 1)
df_inter_gencodeLncRna <- read.table(".temp2/gencode20+.intersecting.master.sig", header=F, sep="\t") %>%
  rename(icID = "V1") %>% 
  mutate(gencodeLncRna = 1)
df_inter_H_concat_catalogs <- read.table(".temp2/H_concat_catalogs.intersecting.master.sig", header=F, sep="\t") %>%
  rename(icID = "V1") %>%
  mutate(H_concat_catalogs = 1)
df_inter_lncBookv2_0 <- read.table(".temp2/lncRNA_LncBookv2_0_GRCh38.intersecting.master.sig", header=F, sep="\t") %>%
  rename(icID = "V1") %>%
  mutate(lncBookv2_0 = 1)
df_inter_mitrans <- read.table(".temp2/miTranscriptome.intersecting.master.sig", header=F, sep="\t") %>%
  rename(icID = "V1") %>% 
  mutate(mitrans = 1)
df_inter_noncode <- read.table(".temp2/NONCODE.intersecting.master.sig", header=F, sep="\t") %>%
  rename(icID = "V1") %>% 
  mutate(noncode = 1)
df_inter_refseq <- read.table(".temp2/refSeq.intersecting.master.sig", header=F, sep="\t") %>%
  rename(icID = "V1") %>% 
  mutate(refseq = 1)

# Join tables
df_masterIDs %>% 
  left_join(x = .,
            y = df_inter_bigtrans,
            by = "icID") %>% 
  left_join(x = .,
            y = df_inter_chess3_1_2,
            by = "icID") %>%
  left_join(x = .,
            y = df_inter_fantomCat,
            by = "icID") %>% 
  left_join(x = .,
            y = df_inter_gen27,
            by = "icID") %>% 
  left_join(x = .,
            y = df_inter_gen47,
            by = "icID") %>% 
  left_join(x = .,
            y = df_inter_gencodeLncRna,
            by = "icID") %>% 
  left_join(x = .,
            y = df_inter_H_concat_catalogs,
            by = "icID") %>%
  left_join(x = .,
            y = df_inter_lncBookv2_0,
            by = "icID") %>%
  left_join(x = .,
            y = df_inter_mitrans,
            by = "icID") %>% 
  left_join(x = .,
            y = df_inter_noncode,
            by = "icID") %>% 
  left_join(x = .,
            y = df_inter_refseq,
            by = "icID") %>% 
  # Exclude master column
  select(!master) %>% 
  # Convert NA to 0
  mutate(across(where(is.numeric), ~ifelse(is.na(.x), 0, .x))) -> df_one_hot_encoded

# Prepare labels  
pretty_labels <- c(
  chess3_1_2 = "CHESS 3.1.2",
  lncBookv2_0 = "lncBook v2.0",
  bigtrans = "bigTranscriptome",
  gencodeLncRna = "GENCODEv20+",
  refseq = "refSeq",
  fantomCat = "fantomCat",
  mitrans = "miTranscriptome",
  noncode = "NONCODE",
  H_concat_catalogs = "lncRNA-merge",
  gen27 = "GENCODE v27",
  gen47 = "GENCODE v47"
)

# Create Upset plot - with lncRNA-merge
df_one_hot_encoded %>% 
  as.data.table() %>%
  setnames(old = names(pretty_labels), new = pretty_labels) %>%
  upset(., 
        sets = pretty_labels,
        mb.ratio = c(0.55, 0.45),
        keep.order = TRUE,
        order.by = "freq",
        set_size.show = TRUE,
        set_size.numbers_size = 4) -> upSet_plot

# Preview
upSet_plot
```

# Save plot to pdf - with lncRNA-merge
```{r}
pdf("IC_upSet_plot.pdf", bg = "white", width=14, height=9)
upSet_plot
```




